### 2024-04-05 universities.Rmd
maps code

```{r}
# read in packages and datasets
install.packages("usmap")
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization
```

```{r}
# chloropleth map visualization

# Carnegie classification data 
universities <- read_excel("cc.xlsx")

# create new dataframe 
univ_and_state <- select(universities, name, state)
head(univ_and_state)

# change column names in jobdata so it is the same as cc data set
names(jobdata)[6] <- "University Name"
names(univ_and_state)[1] <- "University Name"

# remove () from jobdata
jobdata$`University Name` <-  gsub(r"{\s*\([^\)]+\)}","",as.character(jobdata$`University Name`))

# merge datasets according to values in University Name
merged <- merge(jobdata, univ_and_state, "University Name") # only loading 264 universities of 553 obs.

# sum of occurrences of ads per each university 
counts <- data.frame(table(jobdata$`University Name`))
names(counts)[1] <- "University Name"
counts_per_state <- merge(counts, univ_and_state, "University Name")

# sum per state
state_counts <- data.frame(table(merged$state))
names(state_counts)[1] <- "state"

# plot freq in ads / state
usmap::plot_usmap(data = state_counts, values = "Freq") +
  labs(title = "Ad Frequency Per Universities in Each State from 2012-2013 to 2022-2023")

ggsave(here("analysis",
            "figures", 
            "freq-by-state-map.png"),
       bg ="white",
       h = 10, # experiment with h and w to get the right size and proportion
       w = 12,
       units = "in",
       dpi = 900) # make the image nice and crisp))
```

```{r}
# get 50 states
all_states <-
all_states
```

```{r}
# counts per state 2012-2023 & 2022-2023

states <- unique(merged$state)
names(states)[1] <- "state"

# 2012 - 2013 subset
subset_2012_2013 <- merged %>%
  filter(year_ad_posted == "2012-2013")

subset_2022_2023 <- merged %>%
  filter(year_ad_posted == "2022-2023")

state_counts_2012_2013 <- 
  data.frame(table(subset_2012_2013$state))

state_counts_2022_2023 <- data.frame(table(subset_2022_2023$state))

# rename to specify year

names(state_counts_2012_2013)[1] <- "state"
names(state_counts_2012_2013)[2] <- "freq_2012_2013"

names(state_counts_2022_2023)[1] <- 
  "state"
names(state_counts_2022_2023)[2] <-
  "freq_2022_2023"

# merge all states and count 
x <- merge(states, state_counts_2012_2013, "state", all = TRUE)
x[is.na(x)] <- 0

merged_state_counts <- merge(x, state_counts_2022_2023, "state", all = TRUE)
merged_state_counts[is.na(merged_state_counts)] <- 0

# create new column

merged_state_counts$perc_change <- (merged_state_counts$freq_2022_2023 / merged_state_counts$freq_2012_2013)

merged_state_counts$Change <- ifelse(merged_state_counts$freq_2012_2013 < merged_state_counts$freq_2022_2023, 1, -1)

# help with if value stays the same set equql to 0, set as =< ?

#merged_state_counts$Change <- ifelse(merged_state_counts$freq_2012_2013 =< merged_state_counts$freq_2022_2023, 1, -1)

# set states with no change to 0
merged_state_counts$Change[is.na(merged_state_counts$Change)] <- 0

# create map of change 
usmap::plot_usmap(data = merged_state_counts, values = "Change") +
  labs(title = "Increase or Decrease in ad Counts per State from 2012-2013 to 2022-2023")

ggsave(here("analysis",
            "figures", 
            "change-map.png"),
       bg ="white",
       h = 10, # experiment with h and w to get the right size and proportion
       w = 12,
       units = "in",
       dpi = 900) # make the image nice and crisp))

```