### 2025-07-09 README.Rmd
update readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/main?urlpath=rstudio) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15847809.svg)](https://doi.org/10.5281/zenodo.15847809)

This repository contains the data and code for our paper:

> Ben Marwick, Anne Marie Poole, Ailin Zhang, Setareh Shafizadeh, and Jess Beck (2025). Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology. _American Antiquity_

A pre-print of this paper is freely available on SocArXiv: https://osf.io/preprints/socarxiv/whkjs

An earlier version was presented as a poster at the Society of American Archaeology Annual Meeting in New Orleans 17-21 April 2024:

> Poole, Anne Marie, Ailin Zhang, Setareh Shafizadeh, Jess Beck, Ben Marwick, 2024. Careers in Ruins: Academic Archaeology Job Trends From 2013 – 2023. Poster presented at the SAA annual meeting in New Orleans, 17-21 April 2024. doi: 10.17605/OSF.IO/J58KH

![](README-fig-001.png)

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): PDF file of the poster we presented at the SAA. Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. 
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `renv::restore()` to ensure you have the packages this analysis depends on 
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.